\vspace{-1in} Demystifying the Computer Science \\PhD Admission in US Universities\\{\large A Guide for International Students

### 2023-10-24 demystify.tex
updates

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? \textbf{Vu}: see sec ~\ref{sec:contact} to contact a professor for research opportunities.
    \end{commentbox}

@@

      \begin{center}
        \includegraphics[scale=0.5]{c5.png}
      \end{center}

If you're reaching out to a professor for the first time,  address them as either Prof. Lastname or Dr. Lastname. I've noticed that many international students tend to use the titles Prof. or Dr. FirstName LastName, but this can come across as if you're simply copying and pasting names. It's not necessary, so stick with Prof. or Dr. Lastname.

Furthermore, it's important to avoid using Mr. or Mrs., or the professor's first name if you're not acquainted with them yet. This approach may be seen as somewhat disrespectful, especially in academic fields where most professors hold advanced degrees. As you become more familiar with a professor and depending on their preferences, you may transition to addressing them by their first name. For instance, my students and colleagues call me Vu. Some students also refer to me as Dr. Vu, which I find somewhat amusing but am perfectly fine with.

Note that in some universities, e.g., those that do not offer PhD degrees, the formal title Dr. Lastname is commonly used. It's essential to observe and follow the conventions at your particular institution. Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof. Lastname is a suitable alternative.

\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname (or Dr./Prof., if you want to be formal) as mentioned above. You can also include their institution if it makes it more precise.  For example, I did my postdoc with Jeff Foster at Univ. of Maryland.   Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone. I see many international students include a lengthy title of people they know, e.g., my advisor is Asst. Prof. X, and I also collaborate with NASA Distinguished Scientist Y.  This is \emph{not necessary} and makes it look like you're trying to show off your connections. In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to. More on cultural differences in \S\ref{sec:cultural}.

\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisors get. They also wonder why their offer letters sometime mentioned that their benefits appear better more than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.

>>>>

    In other fields, e.g. Software Engineering, Vietnamese students face challenges in reaching US professors. Do you have any tips for Vietnamese students who want to connect with US professors and work as research assistants? \textbf{Vu}: \S\ref{sec:contact} shows how to contact a professor for research opportunities.
    \end{commentbox}

@@

      \begin{center}
        \includegraphics[scale=0.7]{c5.png}
      \end{center}

If you're reaching out to a professor for the first time,  address them as either Prof. Lastname or Dr. Lastname. Many international students use Prof. or Dr. FirstName LastName, but this can come across as if you're simply copying and pasting names. It's not necessary, so stick with Prof. or Dr. Lastname.

Furthermore, avoid using Mr. or Mrs., or the professor's first name if you're not acquainted with them yet.  As you become more familiar with a professor and depending on their preferences, you may transition to addressing them by their first name.
For example, my students and colleague call me Vu. Some students call me \emph{Dr. Vu}, which I find a bit amusing but am totally fine with it.

Note that in some universities the formal title Dr. Lastname is preferred. You just need to observe and follow the conventions at your particular institution. Additionally, be aware that not all faculty members may hold a Ph.D., in which case using Prof. Lastname is a suitable alternative.

\paragraph{Referring to professors you know} Here, because you are already familiar with these individuals, you can just informally use their firstname as mentioned above (or Dr./Prof., if you want to be formal). You can also include their institution if it makes it more precise.  For example, I did my postdoc with Jeff Foster at Univ. of Maryland.   Do not include ranking (e.g., Assistant, Associate, Scientist, ...) when referring to someone. I see many international students include a lengthy title of people they know, e.g., my advisor is Asst. Prof. X, and I also collaborate with Distinguished Scientist Y.  This is \emph{not necessary} and makes it look like you're trying to show off your connections. In short, these nuances represent some cultural differences that you may encounter and will gradually adapt to. More on cultural differences in \S\ref{sec:cultural}.

\subsection{How much do \emph{you} cost?}\label{sec:ra-cost}
PhD students often ask why their salary is so low compared to ludicrous grants their advisors get. They also wonder why their offer letters sometime mentioned that their benefits appear better than what they actually receive (i.e., stipend).  This section aims to shed some light to these questions.
